---
output:
  word_document: default
---
```{r}
library(data.table)
library(PMCMR)
library(userfriendlyscience)
library(PASWR2)
library(psych)
library(BSDA)
```

---
title: "Homework 4 and 5 combined"
author: "Mitesh Ranmal Jain"
date: "11/16/2019"
---

## Question 1

```{r}
Prob_Type_1 <- (1 - (0.2+0.3))

Prob_Type_2 <- (1 - (0.2 + 0.1 + 0.1 + 0.2 + 0.1))
```
#### Answer:
Probability of Type I error is 0.5.
Probability of type II error is 

## Question 2

$$H_0 : \mu \ge  32 $$
$$H_1 : \mu <  32$$

```{r}

Question2 <- read.csv("Question2.csv")
smoke_break_avg <- mean(Question2$Minutes)
smoke_break_std <- sd(Question2$Minutes)
prev_break_avg = 32
prev_break_sd = 8
n <- nrow(Question2)

t.test(Question2$Minutes, alternative = "less", mu = 32)
```

We reject $H_0$ and accept mean time will be less than 32 minutes. 


```{r}
z <-(smoke_break_avg - prev_break_avg)/(prev_break_sd/sqrt(n))

p_sample <- pnorm(z)

z_critical <- qnorm(0.5)

x_critical <- 32 + z_critical*prev_break_sd/sqrt(n)

Type_2_Error_Prob <- 1 - pnorm(x_critical, mean = 30, prev_break_sd/sqrt(n))
```
#### Answer:

## Question 3

$\mu_0 : Mean\ of\ sales\ at\  \$9$
$\mu_1 : Mean\ of\ sales\ at\  \$10$
$\mu_2 : Mean\ of\ sales\ at\  \$11$

$$H_0 : \mu_0 = \mu_1 = \mu_2$$
$$H_1 : Not\ all\ \mu\ are\ equal$$

```{r}

mean_sales_9 =  153.60
sd_sales_9 = 25.57
n_sales_9 = 20

mean_sales_10 = 151.50
sd_sales_10 = 30.39
n_sales_10 = 20

mean_sales_11 = 133.25
sd_sales_11 = 25.03
n_sales_11 = 20

grand_mean = (mean_sales_9 + mean_sales_10 + mean_sales_11)/3
total_n = 60

sst = (20*((mean_sales_9 - grand_mean)^2)) + (20*((mean_sales_10 - grand_mean)^2)) + (20*((mean_sales_11 - grand_mean)^2))

sse = (19*((sd_sales_9)^2)) + (19*((sd_sales_10)^2)) + (19*((sd_sales_11)^2))

mst = sst/2

mse = sse/(60-3)

f_statistic = mst/mse

f_critical = qf(p = 0.05, df1 =  2, df2 = 57, lower.tail = FALSE)

if(f_statistic > f_critical){
  print("The manager should conclude that the sales volume WILL differ significantly if the new product is priced $9 or $11")
} else {
  print("The manager should conclude that the sales volume WILL NOT differ significantly if the new product is priced $9 or $11")
}
```

$$H_0 : \mu_1 = \mu_2 $$
$$H_1 : \mu\ are\ not\ equal$$
```{r}
grand_mean = (mean_sales_9 + mean_sales_10)/2
total_n = 40

sst = (20*((mean_sales_9 - grand_mean)^2)) + (20*((mean_sales_10 - grand_mean)^2))

sse = (19*((sd_sales_9)^2)) + (19*((sd_sales_10)^2))

mst = sst/2

mse = sse/(40-2)

f_statistic = mst/mse

f_critical = qf(p = 0.05, df1 =  1, df2 = 38, lower.tail = FALSE)

if(f_statistic > f_critical){
  print("The manager should conclude that the sales volume WILL differ significantly if the new product is priced $9.")
} else {
  print("The manager should conclude that the sales volume WILL NOT differ significantly if the new product is priced $9.")
}
```
$$H_0 : \mu_2 = \mu_3 $$
$$H_1 : \mu\ are\ not\ equal$$
```{r}
grand_mean = (mean_sales_11 + mean_sales_10)/2
total_n = 40

sst = (20*((mean_sales_11 - grand_mean)^2)) + (20*((mean_sales_10 - grand_mean)^2))

sse = (19*((sd_sales_11)^2)) + (19*((sd_sales_10)^2))

mst = sst/2

mse = sse/(40-2)

f_statistic = mst/mse

P_10_11 <- pf(f_statistic, 1, 38, lower.tail = FALSE)

if(P_10_11 > 0.05){
  print("The manager should conclude that the sales volume WILL differ significantly if the new product is priced $11.")
} else {
  print("The manager should conclude that the sales volume WILL NOT differ significantly if the new product is priced $11.")
}
```
#### Answer:
The sales volume will differ significantly if the product is priced at $11.

## Question 4

```{r}

Question4 <- read.csv("Question4.csv")

Cereal <- factor(Question4$Cereal)
Age <- Question4$Age
Income <- Question4$Income
Education <- Question4$Education
```

$\mu_1 : Mean\ of\ cereal\ 1$
$\mu_2 : Mean\ of\ cereal\ 2$
$\mu_3 : Mean\ of\ cereal\ 3$
$\mu_4 : Mean\ of\ cereal\ 4$


### Part a.

$$H_0 : \mu_1 = \mu_2 = \mu_3 = \mu_4$$
$$H_1 : Not\ all\ \mu\ are\ equal$$

```{r}

model_4_a <- lm(Age ~ Cereal)
residual_4_a <- resid(model_4_a)
predict_4_a <- predict(model_4_a)

## Test for normality

nortest::ad.test(residual_4_a)

hist(residual_4_a)
```

```{r}

## Test for equal variances

bartlett.test(Age ~ Cereal)

boxplot(Age ~ Cereal)
```

```{r}

oneway.test(Age ~ Cereal, var.equal = FALSE)
```

p-value is less than 0.05, hence we can conclude that there are differences between the ages of the consumers of the 4 different cereals.

```{r}

posthocTGH(Age, Cereal, method = "games-howell")
```

#### Answer: 
There are differences between the ages of the consumers of the four cereals 

### Part b.

$$H_0 : \mu_1 = \mu_2 = \mu_3 = \mu_4$$
$$H_1 : Not\ all\ \mu\ are\ equal$$

```{r}

model_4_b <- lm(Income ~ Cereal)
residual_4_b <- resid(model_4_b)
predict_4_b <- predict(model_4_b)

## Test for normality

nortest::ad.test(residual_4_b)

hist(residual_4_b)
```

```{r}

## Test for equal variances

bartlett.test(Income ~ Cereal)

boxplot(Income ~ Cereal)
```

```{r}

ANOVA_4_b <- aov(Income ~ Cereal)
summary(ANOVA_4_b)
```
p-value is less than 0.05, hence we can conclude that there are differences between the incomes of the consumers of the 4 different cereals.
```{r}
plot(TukeyHSD(ANOVA_4_b), las = 1)
```

#### Answer: 
There are differences between the incomes of the consumers of the four cereals.

### Part c.

$$H_0 : \mu_1 = \mu_2 = \mu_3 = \mu_4$$
$$H_1 : Not\ all\ \mu\ are\ equal$$

```{r}

model_4_c <- lm(Education ~ Cereal)
residual_4_c <- resid(model_4_c)
predict_4_c <- predict(model_4_c)

## Test for normality

nortest::ad.test(residual_4_c)

hist(residual_4_c)
```

```{r}

## Test for equal variances

bartlett.test(Education ~ Cereal)

boxplot(Education ~ Cereal)
```

```{r}

ANOVA_4_c <- aov(Education ~ Cereal)
summary(ANOVA_4_c)
```
```{r}
plot(TukeyHSD(ANOVA_4_c), las = 1)
```

#### Answer: 
There are NO differences between the education of the consumers of the four cereals.

### Part d.

```{r}

```

## Question 5
$\mu_1 : Mean\ of\ diet\ 1$
$\mu_2 : Mean\ of\ diet\ 2$
$\mu_3 : Mean\ of\ diet\ 3$
$\mu_4 : Mean\ of\ diet\ 4$

$$H_0 : All\ means\ are\ equal$$
$$H_1 : Atleast\ two\ means\ differ$$

```{r}

Question5 <- read.csv("Question5.csv")

Group <- factor(Question5$Group)
Diet <- factor(Question5$Treatment)
Weight_Loss <- Question5$WeightLoss

model_5 <- lm(Weight_Loss ~ Group + Diet)
resids_5 <- residuals(model_5)
predict_5 <- predict(model_5)

## Normality Tests

nortest::ad.test(resids_5)

qqnorm(resids_5)
qqline(resids_5)
```

```{r}
## Variance Tests

plot(resids_5 ~ predict_5)

car::leveneTest(Weight_Loss ~ Diet)

car::leveneTest(Weight_Loss ~ Group)

```
```{r}
ANOVA_5 <- aov(Weight_Loss ~ Group + Diet)
summary(ANOVA_5)
```

```{r}
TukeyHSD(ANOVA_5, which = 'Diet', ordered = TRUE, conf.level = 0.99)
plot(TukeyHSD(ANOVA_5, which = 'Diet', ordered = TRUE, conf.level = 0.99))

```
### Answer:

Randomized Block Design Model was used for this experiment.
The p-values for the treatment(Diet) and blocks(Group) is NOT statitistically significant at 1% confidence level. So we can conclude that there is NO difference among the 4 diets and blocking was NOT meaningful.
From this we can conclude that the experimental design was NOT sound.

## Question 6

```{r}
Question6 <- read.csv("Question6.csv")

stack_6_business <- Question6[, c(1,2)]
stack_6_arts <- Question6[, c(1,3)]
stack_6_science <- Question6[, c(1,4)]

stack_6_business["Degree"] = "Business"
stack_6_arts["Degree"] = "Arts"
stack_6_science["Degree"] = "Science"

names(stack_6_arts)[2] <- "Commission"
names(stack_6_business)[2] <- "Commission"
names(stack_6_science)[2] <- "Commission"

stack_6 <- rbind(stack_6_arts, stack_6_business, stack_6_science)

Group_6 <- factor(stack_6$Group)
Commission <- stack_6$Commission
Degree <- factor(stack_6$Degree)
```

#### Part a.

$$H_0 : All\ means\ are\ equal$$
$$H_1 : Atleast\ two\ means\ differ$$

```{r}
model_6 <- lm(Commission ~ Group_6 + Degree)
resids_6 <- residuals(model_6)
predict_6 <- predict(model_6)

## Normality Tests

nortest::ad.test(resids_6)
```
```{r}
hist(resids_6)
```

```{r}
## Variance Tests

plot(resids_6 ~ predict_6)
```
```{r}
car::leveneTest(Commission ~ Degree)
```
```{r}
car::leveneTest(Commission ~ Group_6)
```

```{r}
Anova_6 <- aov(Commission ~ Group_6 + Degree)
summary(Anova_6)
```

```{r}
TukeyHSD(Anova_6, which = 'Degree', ordered = TRUE)
plot(TukeyHSD(Anova_6, which = 'Degree', ordered = TRUE))
```
#### Answer:

#### Part b.

$$H_0 : All\ means\ are\ equal$$
$$H_1 : Atleast\ two\ means\ differ$$

```{r}
model_6_degree  <- lm(Commission ~ Degree)
resids_6_degree <- residuals(model_6_degree)
predict_6_degree <- predict(model_6_degree)

## Normality Tests

nortest::ad.test(resids_6_degree)
```
```{r}
hist(resids_6_degree)
```

```{r}
## Variance Tests

plot(resids_6_degree ~ predict_6_degree)
```
```{r}
car::leveneTest(Commission ~ Degree)
```

```{r}
Anova_6_Degree <- aov(Commission ~ Degree)
summary(Anova_6_Degree)
```
```{r}
TukeyHSD(Anova_6_Degree, which = 'Degree', ordered = TRUE)
plot(TukeyHSD(Anova_6_Degree, which = 'Degree', ordered = TRUE))
```
#### Answer: 
If only degree is taken into account there is NO sufficient evidence to allow the recruiter to conclude that there are differences in sales ability between the holders of the three types of degrees.

## Question 7

### Test for differences between the levels of Age

$$H_0 : Age\ has\ no\ effect\ on\ miles\ driven $$
$$H_1 : Age\ has\ effect\ on\ miles\ driven $$

### Test for differences between the levels of Gender

$$H_0 : Gender\ has\ no\ effect\ on\ miles\ driven $$
$$H_1 : Gender\ has\ effect\ on\ miles\ driven $$

### Test for interaction between Age and Gender

$$H_0 : Age\ and\ Gender\ do\ not\ interact\ to\ affect\ miles\ driven $$
$$H_1 : Age\ and\ Gender\ do\ interact\ to\ affect\ miles\ driven $$

```{r}

Question7 <- read.csv("Question7.csv")

test <- Question7
Question7[1:20, 1] = "Males"
Question7[21:40, 1] = "Females"

Age1 <- Question7[, c(1,2)]
Age1["Age"] = "Age1"
names(Age1)[2] = "MilesDriven"
names(Age1)[1] = "Gender"

Age2 <- Question7[, c(1,3)]
Age2["Age"] = "Age2"
names(Age2)[2] = "MilesDriven"
names(Age2)[1] = "Gender"

Age3 <- Question7[, c(1,4)]
Age3["Age"] = "Age3"
names(Age3)[2] = "MilesDriven"
names(Age3)[1] = "Gender"

Age4 <- Question7[, c(1,5)]
Age4["Age"] = "Age4"
names(Age4)[2] = "MilesDriven"
names(Age4)[1] = "Gender"

Age5 <- Question7[, c(1,6)]
Age5["Age"] = "Age5"
names(Age5)[2] = "MilesDriven"
names(Age5)[1] = "Gender"

stack_7 <- rbind(Age1, Age2, Age3, Age4, Age5)

Gender <- factor(stack_7$Gender)
MilesDriven <- as.numeric(stack_7$MilesDriven)
Age_7 <- factor(stack_7$Age)

model_7 <- lm(MilesDriven ~ Gender + Age_7 + Gender*Age_7)
resids_7 <- residuals(model_7)
predicts_7 <- predict(model_7)

## Test for normality

nortest::ad.test(resids_7)
shapiro.test(resids_7)

hist(resids_7)

## The data is normal
```
```{r}
## Variance test

car::leveneTest(MilesDriven ~ Gender)
car::leveneTest(MilesDriven ~ Age_7)
```
```{r}
## ANOVA

ANOVA_7 <- aov(MilesDriven ~ Gender + Age_7 + Gender*Age_7)
summary(ANOVA_7)

interaction.plot(Age_7, Gender,  MilesDriven)
```
#### Answer:
We do NOT have enough evidence to conclude that males and female drivers differ in the number of miles they drive.
We CAN infer that there are differences between the age categories in the number of miles they drive


## Question 8

$$H_0 : p_1 = 0.105,\ p_2 = 0.219,\ p_3 =0.533,\ p_4 = 0.143 $$
$$H_1 : Atleast\ one\ proportion\ differs\ from\ the\ proposed\ plan$$

```{r}
Observed_8 = c(0,20,83,52)
Prob_8 = c(0.105,0.219,0.533,0.143)

chisq.test(x = Observed_8,p = Prob_8)
```
#### Answer:
At least 1 frequency after the contest differs from the proportions before the contest.

## Question 9

$H_0$: Market Direction and Day of the week are independent. 
$H_1$: Market Direction and Day of the week are not independent.

```{r}
stack_9_data <- data.frame(c(rep(c('Down'),5),
                             rep(c('Up'),5)), 
                           rep(c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'),2),
                           c(42,49,46,43,41,
                             53,55,58,59,58))

colnames(stack_9_data) <- c("Flux", "Day", "Return")

stack_9_data <- stack_9_data[rep(seq_len(nrow(stack_9_data)), stack_9_data$Return),]

chisq.test(stack_9_data$Flux, stack_9_data$Day)

mosaicplot(~ Day + Flux, data = stack_9_data,
            main = "Mosiac Level", xlab = "Day", ylab = "Flux",
           col = c (18,3))

```
#### Answer:
p-value is greater than 0.05 and we fail to reject $H_0$ i.e. Market Direction and Day of the week are independent.

## Question 10
$H_0$: The two population locations are the same

$H_1$: The location of the assets of households with college degrees is to the right of the location of assets of households with some college

```{r}
Question10 <- read.csv("Question10.csv")

Question10$EDCL <- factor(Question10$EDCL)

Question10 <- Question10[Question10$EDCL == "4"|Question10$EDCL == "3",]

EDCL <- factor(Question10$EDCL)
ASSETS <- Question10$ASSET

ASSETS <- as.numeric(gsub(",", "", ASSETS))

boxplot(ASSETS ~ EDCL)
```

```{r}
tapply(ASSETS, EDCL, sd)

model_10 <- lm(ASSETS ~ EDCL)
resids_10 <- residuals(model_10)
preds_10 <- predict(model_10)

nortest::ad.test(resids_10)

hist(resids_10)

car::leveneTest(ASSETS ~ EDCL)
```

```{r}
### Transform using log

Transformed_ASSETS <- log(ASSETS)

model_10_transformed <- lm(Transformed_ASSETS ~ EDCL)
resids_10_transformed <- residuals(model_10_transformed)
preds_10_transformed <- predict(model_10_transformed)

nortest::ad.test(resids_10_transformed)

## The data is not normally distributed

hist(resids_10_transformed)
```

```{r}
## Test for variance

car::leveneTest(Transformed_ASSETS ~ EDCL)

## The variances are almost equal

wilcox.test(Transformed_ASSETS ~ EDCL, alt = "less", paired = FALSE)
```

### Answer:
There is enough evidence to conclude that heads of households with college degrees have more assets than those who have some college i.e. we reject $H_0$.

## Question 11

$$H_0 : All\ means\ are\ equal$$
$$H_1 : All\ means\ differ$$

```{r}
Question11 <- read.csv("Question11.csv")
stacked_11 <- stack(Question11)

wait_time <- stacked_11$values
campus <- stacked_11$ind

model_11 <- lm(wait_time ~ campus)
resids_11 <- residuals(model_11)
predicts_11 <- predict(model_11)

## Normality Test

nortest::ad.test(resids_11)

hist(resids_11)
```

```{r}
## Test for constant variance

bartlett.test(wait_time ~ campus)

boxplot(wait_time ~ campus)
```
```{r}
ANOVA_11 <- aov(wait_time ~ campus)
summary(ANOVA_11)

## There is significant difference in the mean wait time at the different campuses

posthocTGH(wait_time, campus, method = "games-howell")

TukeyHSD(ANOVA_11, las = 1)
plot(TukeyHSD(ANOVA_11, las = 1))
```

#### Answer:
There is evidence of a difference in waiting time.

## Question 12
$\mu_1 : Mean\ weight-loss\ of\ diet\ 1$
$\mu_2 : Mean\ weight-loss\ of\ diet\ 2$
$\mu_3 : Mean\ weight-loss\ of\ diet\ 3$

$$H_0 : All\ means\ are\ equal$$
$$H_1 : Atleast\ two\ mwans\ differ$$

```{r}
Question12 <- read.csv("Question12.csv")

stack_12 <- melt(data = Question12, id.vars = c("Person"), measure.vars = c("Original", "New.Recipe.1", "New.Recipe..2"))

Rating <- stack_12$value
Person <- factor(stack_12$Person)
Recipe <- factor(stack_12$variable)

model_12 <- lm(Rating ~ Recipe)
resids_12 <- residuals(model_12)
predicts_12 <- predict(model_12)

nortest::ad.test(resids_12)

## The data is normal
```

```{r}
## Check for variance
bartlett.test(Rating ~ Recipe)

## The data satisies constant variance requirement
```
```{r}

ANOVA_12 <- aov(Rating ~ Recipe + Person)
summary(ANOVA_12)

## There is significant difference in ratings at 5% significance level
```
```{r}
TukeyHSD(ANOVA_12, which = 'Recipe', ordered = TRUE)
```
```{r}
plot(TukeyHSD(ANOVA_12, which = 'Recipe', ordered = TRUE))
```

#### Answer:
The original recipe and the New Recipe 2 differ in ratings.

## Question 13


$$H_0 : All\ means\ are\ equal$$
$$H_1 : Atleast\ two\ means\ differ$$

```{r}
Question13 <- read.csv("Question13.csv")

## Test for normality

stack_13 <- melt(Question13, id.vars = c("Year"), measure.vars = c("Growth", "Value"))

Year <- stack_13$Year
Investment <- stack_13$variable
Annual_Return <- stack_13$value

model_13 <- lm(Annual_Return ~ Year + Investment)
resids_13 <- residuals(model_13)
predicts_13 <- predict(model_13)

hist(resids_13)

nortest::ad.test(resids_13)

## The data is normally distributed
```

```{r}
## Test for variance

bartlett.test(Annual_Return ~ Investment)

## The test is satisfied
```

```{r}
cor.test( as.numeric(Investment), as.numeric(Annual_Return))
```

#### Answer:
The performance of the returns of these two funds were NOT similar over the 2007-2016 timeframe.

## Question 14

$$H_0 : \rho\ = 0$$
$$H_1 : \rho\ \neq 0$$

```{r}
Question14 <- read.csv("Question14.csv")

cor.test(x  = Question14$JOBLOSE, y = Question14$HRS1, method = "spearman")
```
#### Answer:
There is a weak positive correlation between long hours worked and chances of losing one's job.

## Question 15

$$H_0 : The\ two\ brands'\ ratings\ are\ the\ same$$
$$H_1 : The\ ratings\ of\ European\ Brand\ is\ greater\ than\ ratings\ of\ Domestic\ Brand$$

```{r}
Question15 <- read.csv("Question15.csv")

SIGN.test(x = Question15$European,
          y = Question15$Domestic,
          alternative = "greater",
          conf.level = 0.90)
```
#### Answer:
The data provide sufficient evidence to infer that the European brand is perceived to be more effective than Domestic brand as the p-value is less than 0.90.

## Question 16

$$H_0 : All\ means\ are\ equal$$
$$H_1 : Atleast\ two\ means\ differ$$

```{r}
Question16 <- read.csv("Question16.csv")
 
stack__16 <- melt(data = Question16, id.vars = c("Title"), measure.vars = c("Amazon", "BN"))

Title <- stack__16$Title
Retailer <- stack__16$variable
Cost <- stack__16$value

model_16 <- lm(Cost ~ Retailer + Title)
resids_16 <- residuals(model_16)
predict_16 <- predict(model_16)

## Normality Tests

nortest::ad.test(resids_16)
shapiro.test(resids_16)

qqnorm(resids_16)
qqline(resids_16)

hist(resids_16)

## Variance Tests

bartlett.test(Cost ~ Retailer)

car::leveneTest(Cost ~ Retailer)

### Constant variance requirement satisfied for Diet

ANOVA_16 <- aov(Cost ~ Retailer + Title)
summary(ANOVA_16)

TukeyHSD(ANOVA_16, which = 'Retailer', ordered = TRUE)
```
#### Answer:
The ð‘-values for both Retailer and Title are statistically significant. So, we conclude that there is a difference among the 2 retailers AND blocking was meaningful.

## Question 17

$H_0$: The data comes from a normal distribution.
$H_1$: The data does not follow a normal distribution.

```{r}
observed_17 <- c(10, 18, 48, 16, 8)
prob_1 <- pnorm(-1.5)
prob_2 <- pnorm(-0.5) - pnorm(-1.5)
prob_3 <- pnorm(0.5) - pnorm(-0.5)
prob_4 <- pnorm(1.5) - pnorm(0.5)
prob_5 <- pnorm(1.5, lower.tail = FALSE)

probability_17 <- c(prob_1, prob_2, prob_3, prob_4, prob_5)

chisq.test(x = observed_17, p = probability_17)
```

#### Answer:
As p-value is less than 0.05, we fail to reject $H_0$ i.e. we can conclude that the data comes from a normal distribution.