---
title: "Statistical Exploration and Reasoning Assignment 3"
author: "Mitesh Ranmal Jain"
date: "10/21/2019"
output: word_document
---
## Question 1

$$\mu_1 = Average\ rating\ of\ exercisers$$
$$\mu_2 = Average\ rating\ of\ non-exercisers$$


$$H_0 : \mu_1 - \mu_2 \leq D_0$$
$$H_1 : \mu_1 - \mu_2 > D_0$$

### Part a.

```{r}
question_1 <- read.csv('Question 1.csv')
exercisers <- question_1[which(question_1$Exerciser == 'Yes'), ]
non_exercisers <- question_1[which(question_1$Exerciser == 'No'), ]

var.test(exercisers$Rating, non_exercisers$Rating, ratio = 1, alternative = 'two.sided')
```
Variances are equal, so run Equal Variance t-test

```{r}
t.test(exercisers$Rating, non_exercisers$Rating, alternative = "greater", mu = 0, paired = FALSE, var.equal = TRUE)
```
$p-value(0.009711)\ is\ less\ than\ \alpha(0.05),\ therefore\ reject\ null\ hypothesis$

### Part b.

$Company\ can\ say\ that\ exercisers\ outperform\ non-exercisers$

## Question 2

$$\mu_1 = Average\ appraised\ value$$
$$\mu_2 = Average\ selling\ price$$

$$H_0 : \mu_1 - \mu_2 \leq D_0$$
$$H_1 : \mu_1 - \mu_2 > D_0$$

```{r}
question_2 <- read.csv('Question 2.csv')
```

$$\alpha = 0.05$$

```{r}
var.test(question_2$Value, question_2$Price, ratio = 1, alternative = 'two.sided')
```
Variances are not equal, so run Unequal Variance t-test
```{r}
t.test(question_2$Value, question_2$Price, alternative = 'two.sided', mu = 0, paired = TRUE, var.equal = FALSE)
```
$p-value(0.7236)\ is\ greater\ than\ \alpha(0.05),\ therefore\ do\ not\ reject\ null\ hypothesis$

$$\alpha = 0.01$$

```{r}
var.test(question_2$Value, question_2$Price, ratio = 1, alternative = 'two.sided', conf.level = 0.99)
```
Variances are not equal, so run Unequal Variance t-test

```{r}
t.test(question_2$Value, question_2$Price, alternative = 'two.sided', mu = 0, paired = TRUE, var.equal = FALSE, conf.level = 0.99)
```
p-value(0.7236) is greater than 0.01, therefore do not reject null hypothesis

$$\alpha = 0.1$$

```{r}
var.test(question_2$Value, question_2$Price, ratio = 1, alternative = 'two.sided', conf.level = 0.9)
```
Variances are equal, so run equal Variance t-test

```{r}
t.test(question_2$Value, question_2$Price, alternative = 'two.sided', mu = 0, paired = TRUE, var.equal = TRUE, conf.level = 0.9)
```
$p-value(0.7236)\ is\ greater\ than\ \alpha(0.01),\ therefore\ do\ not\ reject\ null\ hypothesis$

## Question 3

$$\sigma_1 = Variance\ in\ service\ time\ for\ Teller\ 1$$
$$\sigma_2 = Variance\ in\ service\ time\ for\ Teller\ 2$$

$$H0 : \sigma_1/\sigma_2 = 1$$
$$H1 : \sigma_1/\sigma_2 \neq 1$$

```{r}
question_3 <- read.csv('Question 3.csv')

var.test(question_3$Teller1, question_3$Teller2, ratio = 1, alternative = 'two.sided', conf.level = 0.9)
```
$p-value\ is\ 1.045e^{-08},\ which\ means\ it\ is\ statistically\ significant\ so\ reject\ H_0$
$Therefore\ the\ variance\ in\ service\ times\ differs\ between\ the\ 2\ tellers$

## Question 4

$$p_1 = Proportion\ of\ people\ who\ took\ Vioxx\ and\ developed\ heart\ problems$$
$$p_2 = Proportion\ of\ people\ who\ took\ placebos\ and\ developed\ heart\ problems$$

$$H_0: p_1 - p_2 \leq 0$$
$$H_1: p_1 - p_2 > 0$$

```{r}
prop.test(c(45, 25), c(1287, 1299), alternative = 'greater', correct = FALSE)
```

```{r}
(pop_vioxx = 2000000*(0.03496503))
```

Population  = 69931

```{r}
(pop_not_vioxx = 2000000*(0.01924557))
```

$Population that got heart problems by taking placebos = 38492$

## Question 5

$$p_1 = Proportion\ of\ retention\ after\ getting\ health\ benefits$$
$$p_2 = Proportion\ of\ retention\ after\ getting\ vacation\ benefits$$

$$H_0: p_1 - p_2 \geq 0.05$$
$$H_1: p_1 - p_2 < 0.05$$

```{r}
question_5 <- read.csv('Question 5.csv')
table(question_5)
```

### Part a.

### Part b.
```{r}
p_health_hat = 107/125
p_vacation_hat = 109/140
denom = sqrt(((p_health_hat)*(1 - p_health_hat)/125) + ((p_vacation_hat)*(1 - p_vacation_hat)/140))

z_5_b = ((p_health_hat - p_vacation_hat) - 0.05)/ denom
pnorm(z_5_b, lower.tail = FALSE)
```
$p-value\ is\ 0.28\ which\ is\ greater\ than\ \alpha(0.05), therefore statistically not significant\\Hence\ faliure\ to\ reject\ the\ null\ hypothesis$
### Part c.
```{r}
prop.test(c(107, 109), c(125, 140), alternative = 'greater', correct = FALSE)
```
$p-value = 0.9475$

## Question 6

```{r}
question_6 <- read.csv('Question 6.csv')
```

### Part a.

$$\mu_1 = Average\ RINCOME\ in\ year\ 2000$$
$$\mu_2 = Average\ RINCOME\ in\ year\ 2008$$

$$H_0 : \mu_1 - \mu_2 \geq 0\ \ i.e.\ income\ does\ not\ increase$$
$$H_1 : \mu1 - \mu2 < 0\ \ i.e.\ income\ does\ increase$$

```{r}
var.test(question_6$RINCOME_2000, question_6$RINCOME_2008, ratio = 1, alternative = 'two.sided')
```

Variances are not equal, so run unequal Variance t-test

```{r}
t.test(question_6$RINCOME_2000, question_6$RINCOME_2008, alternative = 'less', var.equal = FALSE, paired = FALSE)
### 
### 
```
$p-value\ (2.29e^{-16})\ is\ statistically\ significant,\ and\ we\ reject\ the\ null\ hypothesis\\Income\ has\ increased\ between\ 2000\ and\ 2008$

### Part b.

$$\mu_1 = Average\ RINCOME\ in\ year\ 2008$$
$$\mu_2 = Average\ RINCOME\ in\ year\ 2014$$

$$H_0 : \mu_1 - \mu_2 \geq 0\ \ i.e.\ income\ does\ not\ increase$$
$$H_1 : \mu1 - \mu2 < 0 \ \ i.e.\ income\ does\ increase$$

```{r}
var.test(question_6$RINCOME_2008, question_6$RINCOME_2014, ratio = 1, alternative = 'two.sided')
```
Variances are not equal, so run unequal Variance t-test

```{r}
t.test(question_6$RINCOME_2008, question_6$RINCOME_2014, alternative = 'less', var.equal = FALSE, paired = FALSE)
```

p-value(0.002308) is statistically significant, so we reject the null hypothesis

### Part c.
```{r}
cpi_data = readxl::read_xlsx('U.S. CPI Annual.xlsx')

cpi_2000 = cpi_data[which(cpi_data$Year == '2000'), 2 ]
cpi_2008 = cpi_data[which(cpi_data$Year == '2008'), 2 ]
cpi_2014 = cpi_data[which(cpi_data$Year == '2014'), 2 ]
```
$$\mu_1 = Average\ adjusted\ RINCOME\ in\ 2008$$
$$\mu_2 = Average\ RINCOME\ in\ 2008$$

$$H_0 : \mu_1 - \mu_2 \geq 0 \ \ i.e.\ income\ did\ not\ increase$$
$$H_1 : \mu_1 - \mu_2 < 0\ \ i.e.\ income\ increased$$
```{r}
adj_rincome_2008_2000 <- question_6$RINCOME_2000*(215.25500/172.1917)

var.test(adj_rincome_2008_2000, question_6$RINCOME_2008, ratio = 1, alternative = 'two.sided')
```
Variances are not equal, so run unequal Variance t-test

```{r}
t.test(adj_rincome_2008_2000, question_6$RINCOME_2008, alternative = 'less', var.equal = FALSE, paired = FALSE)
```
P-VALUE(0.05485) is not statistically significant, so we cannot reject the null hypothesis

Income has not increased between 2000 and 2008 when inflation is adjusted

### Part d.

$$\mu_1 = Average\ adjusted\ RINCOME\ in\ 2014$$
$$\mu_2 = Average\ RINCOME\ in\ 2014$$


$$H_0 : \mu_1 - \mu_2 \geq 0 \ \ i.e.\ income\ did\ not\ increase$$
$$H_1 : \mu_1 - \mu_2 < 0\ \ i.e.\ income\ increased$$



```{r}
adj_rincome_2014_2008 <- question_6$RINCOME_2008*(236.715/215.25500)

var.test(adj_rincome_2014_2008, question_6$RINCOME_2014, ratio = 1, alternative = 'two.sided')
```
Variances are equal, so run equal Variance t-test

```{r}
t.test(adj_rincome_2014_2008, question_6$RINCOME_2014, alternative = 'less', var.equal = TRUE, paired = FALSE)
```
$P-VALUE(0.4849)\ is\ not\ statistically\ significant,\ so\ we\ cannot\ reject\ the\ null\ hypothesis\\Income\ has\ NOT\ increased\ between\ 2008\ and\ 2014\ when\ inflation\ is\ adjusted$

### Part e.


